---
title: "General crop model performance figures"
format: 
  html: 
    toc: true
    code-fold: true
    code-summary: "Show the code"
    embed-resources: true
    self-contained-math: true

editor: visual
---

::: {.callout-note title="Overview" icon="false"}
This notebook will guide you through the code to plot the results of the general model performance (i.e. not under specific extremes) in R. The following figures will be computed:

-   Figure 1: global assessment of crop model performance (1981-2016) for aggregated crops

-   Supplementary figure 1: global performance table without harvest area weighting

-   Supplementary figures 2-5: global performance maps for each crop (mai, soy, wwh, ri1)

-   Supplementary figures 6-10: production maps for each crop (mai, soy, wwh, ri1, aggregated)

The fully processed and filtered data files, available under GGCMI-validation/data/processed/figure_ready_data are required to produce the figures.
:::

We load the necessary libraries

```{r, warning = FALSE, message = FALSE}
library(tidyverse)   # Includes dplyr, ggplot2, readr, etc.
library(sjPlot)  # to set and configure plotting theme
library(rnaturalearth) # to create base world map with countries
library(viridis) # colorblind color schemes
```

We configure the plotting theme based on the data visualization blog from Cédric Scherer: <https://www.cedricscherer.com/>

```{r, warning = FALSE, message = FALSE}
theme_Scherer <- function (base_size = 12, base_family = "Helvetica") {
  half_line <- base_size/2
  theme(
    line = element_line(color = "black", linewidth = .5,
                        linetype = 1, lineend = "butt"),
    rect = element_rect(fill = "white", color = "black",
                        linewidth = .5, linetype = 1),
    text = element_text(family = base_family, face = "plain",
                        color = "black", size = base_size,
                        lineheight = .9, hjust = .5, vjust = .5,
                        angle = 0, margin = margin(), debug = FALSE),
    axis.line = element_blank(),
    axis.line.x = NULL,
    axis.line.y = NULL,
    axis.text = element_text(size = base_size * 1.1, color = "gray30"),
    axis.text.x = element_text(margin = margin(t = .8 * half_line/2),
                               vjust = 1),
    axis.text.x.top = element_text(margin = margin(b = .8 * half_line/2),
                                   vjust = 0),
    axis.text.y = element_text(margin = margin(r = .8 * half_line/2),
                               hjust = 1),
    axis.text.y.right = element_text(margin = margin(l = .8 * half_line/2),
                                     hjust = 0),
    axis.ticks = element_line(color = "gray30", linewidth = .7),
    axis.ticks.length = unit(half_line / 1.5, "pt"),
    axis.ticks.length.x = NULL,
    axis.ticks.length.x.top = NULL,
    axis.ticks.length.x.bottom = NULL,
    axis.ticks.length.y = NULL,
    axis.ticks.length.y.left = NULL,
    axis.ticks.length.y.right = NULL,
    axis.title.x = element_text(margin = margin(t = half_line),
                                vjust = 1, size = base_size * 1.3,
                                face = "bold"),
    axis.title.x.top = element_text(margin = margin(b = half_line),
                                    vjust = 0),
    axis.title.y = element_text(angle = 90, vjust = 1,
                                margin = margin(r = half_line),
                                size = base_size * 1.3, face = "bold"),
    axis.title.y.right = element_text(angle = -90, vjust = 0,
                                      margin = margin(l = half_line)),
    legend.background = element_rect(color = NA),
    legend.spacing = unit(.4, "cm"),
    legend.spacing.x = NULL,
    legend.spacing.y = NULL,
    legend.margin = margin(.2, .2, .2, .2, "cm"),
    legend.key = element_rect(fill = "gray95", color = "white"),
    legend.key.size = unit(1.2, "lines"),
    legend.key.height = NULL,
    legend.key.width = NULL,
    legend.text = element_text(size = rel(.8)),
    legend.text.align = NULL,
    legend.title = element_text(hjust = 0),
    legend.title.align = NULL,
    legend.position = "right",
    legend.direction = NULL,
    legend.justification = "center",
    legend.box = NULL,
    legend.box.margin = margin(0, 0, 0, 0, "cm"),
    legend.box.background = element_blank(),
    legend.box.spacing = unit(.4, "cm"),
    panel.background = element_rect(fill = "white", color = NA),
    panel.border = element_rect(color = "gray30",
                                fill = NA, linewidth = .7),
    panel.grid.major = element_line(color = "gray90", linewidth = 1),
    panel.grid.minor = element_line(color = "gray90", linewidth = .5,
                                    linetype = "dashed"),
    panel.spacing = unit(base_size, "pt"),
    panel.spacing.x = NULL,
    panel.spacing.y = NULL,
    panel.ontop = FALSE,
    strip.background = element_rect(fill = "white", color = "gray30"),
    strip.text = element_text(color = "black", size = base_size),
    strip.text.x = element_text(margin = margin(t = half_line,
                                                b = half_line)),
    strip.text.y = element_text(angle = -90,
                                margin = margin(l = half_line,
                                                r = half_line)),
    strip.text.y.left = element_text(angle = 90),
    strip.placement = "inside",
    strip.placement.x = NULL,
    strip.placement.y = NULL,
    strip.switch.pad.grid = unit(0.1, "cm"),
    strip.switch.pad.wrap = unit(0.1, "cm"),
    plot.background = element_rect(color = NA),
    plot.title = element_text(size = base_size * 1.8, hjust = .5,
                              vjust = 1, face = "bold",
                              margin = margin(b = half_line * 1.2)),
    plot.title.position = "panel",
    plot.subtitle = element_text(size = base_size * 1.3,
                                 hjust = .5, vjust = 1,
                                 margin = margin(b = half_line * .9)),
    plot.caption = element_text(size = rel(0.9), hjust = 1, vjust = 1,
                                margin = margin(t = half_line * .9)),
    plot.caption.position = "panel",
    plot.tag = element_text(size = rel(1.2), hjust = .5, vjust = .5),
    plot.tag.position = "topleft",
    plot.margin = margin(rep(base_size, 4)),
    complete = TRUE
  )
}
set_theme(theme_Scherer())
```

Then we can load our data:

```{r, warning = FALSE, message = FALSE}
base_path <- "" # Where is the code repo stored?

# List of crops
crops <- c("mai", "wwh", "ri1", "soy", "aggr")

# Read each .rds file into a named list
crop_data <- lapply(crops, function(crop) {
  readRDS(file.path(base_path, paste0("GGCMI-validation/data/processed/figure_ready_data/general_", crop, ".rds")))
})

general_sim_aggr <- readRDS(file.path(base_path, "GGCMI-validation/data/processed/figure_ready_data/general_sim_aggr.rds"))
names(crop_data) <- crops
```

## Figure 1

This figure consists of 3 sub-figures that we manually aligned into one overall figure.

### Global map

First we plot the global map showing Rsquared values based on a grid cell level. The values are computed by comparing the benchmark and the model ensemble median time series for the absolute detrended yield values in each grid cell. We also check the 10 best and worst gridcells.

```{r, warning = FALSE, message = FALSE}
ens_map_aggr <- crop_data[["aggr"]] %>% 
  inner_join(general_sim_aggr %>% filter(model == "ensemble") %>% dplyr::select(lat, lon, year, difftrend_sim, ctr) %>% distinct(), by = c("lat", "lon", "year")) %>%
  group_by(lat, lon) %>% 
 filter(sd(difftrend_obs) > 0, sd(difftrend_sim) > 0) %>% # only gridcells where the standard deviation is bigger than 0
    summarise(rsquared = cor(difftrend_obs, difftrend_sim) ^2,
             ctr = ctr) %>%
distinct()
```

Top 10 best performing gridcells

```{r, warning = FALSE, message = FALSE}
head(ens_map_aggr[order(-ens_map_aggr$rsquared), ], 10) 
```

top 10 worst performing gridcells

```{r, warning = FALSE, message = FALSE}
head(ens_map_aggr[order(ens_map_aggr$rsquared), ], 10) 
```

To make a map we first have to create a base map with countries. We can do this using the `ne_countries` function from the `rnaturalearth` package

```{r, warning = FALSE, message = FALSE}
world_geometries <- ne_countries(scale = "medium", returnclass = "sf") %>% 
  mutate(ctr = iso_a3_eh) %>% # eh needed to get France and Norway as well 
  filter(name_long != "	
Ashmore and Cartier Islands", name_long != "Indian Ocean Territories") %>%  # Get rid of duplicate AUS entries
  dplyr::select(ctr,subregion) %>% 
   filter(ctr != "ATA") %>%
 distinct() 

# Create the base map
base_map <- ggplot() +
  geom_sf(data = world_geometries, fill = "gray", color = "black", size = 0.2) +
scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    panel.background = element_rect(fill = "white"),
   panel.grid.major = element_line(color = "gray80", size = 0.2),  # major gridlines
    panel.grid.minor = element_blank(),  # Remove minor gridlines
  )

base_map
```

Then we can plot the gridcell values on the base map to create the sub-figure

```{r, warning = FALSE, message = FALSE}
map_rsquared <- base_map +
  geom_tile(data = ens_map_aggr, aes(x = lon, y = lat, color = rsquared), size = 0.2) +
scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_color_viridis(name = expression(R^2), option = "viridis", limits = c(0, 0.75),
                      guide = guide_colorbar(barwidth = unit(18, "cm"),  # Increase colorbar width
                                            barheight = unit(1, "cm"),  # Set colorbar height
                                            title.position = "top",  # Position title above the colorbar
                                             title.vjust = 1,  # Add space between title and colorbar
                                              title.hjust = 0.5  # Center the title
                      ))+
  labs(x = "", y = "") +
  theme(legend.position = "top")

map_rsquared

# Save the figure
ggsave(file.path(base_path, "GGCMI-validation/results/main/fig1a_GlobalPerformanceMap.png"), map_rsquared, width = 24, height = 12, dpi = 300)
```

### Latitudinal scatterplot

Then we also want to create a scatterplot showing the model ensemble performance (Rsquared) along the latitudinal axis (y). We use the geom_smooth(method = "gam") to fit a smooth nonlinear trendline.

```{r, warning = FALSE, message = FALSE}
# Generate latitudinal distribution plot
lat_plot <- ggplot(ens_map_aggr, aes(x = lat, y = rsquared)) +
  geom_point(alpha = 0.3, color = "#3288bd") +
  geom_smooth(method = "gam", color = "black", fill = "gray80") +
  labs(x = "Latitude", y = expression(R^2)) + 
  theme(axis.title.x = element_text(size = 25), axis.title.y = element_text(size = 12)) +
  coord_flip() +
  scale_x_continuous(limits = c(-45, 60), breaks = c(-40, -20, 0, 20, 40))
  ylim(0, 1)

lat_plot

# Save the latitudinal plot
ggsave(file.path(base_path, "GGCMI-validation/results/main/fig1b_LatitudinalScatter.png"),
     lat_plot, width = 12, height = 12, dpi = 300)
```

Finally, we also create a heatmap table showing the general global performance for all models across different type of performance metrics: RMSE, Rsquared, KGE. Next the to performance metrics we also show the variability (sd) and we order the models from high to low variability.

We first define a function to calculate RMSE and KGE. Note that for KGE we use a self-customized version (KGE'). The original KGE can be calculated as KGE = 1-KGE'.

```{r, warning = FALSE, message = FALSE}
# Define RMSE function
rmse <- function(observed, predicted) {
  sqrt(mean((observed - predicted) ^ 2))
}

# Function to calculate KGE components
calculate_kge<- function(benchmark, simulated) {
  r <- cor(benchmark, simulated, use = "complete.obs")  # Correlation
  Beta <- (mean(simulated, na.rm = TRUE) - mean(benchmark, na.rm = TRUE)) / sd(benchmark, na.rm = TRUE)  # Bias (Beta)
  Alpha <- sd(simulated, na.rm = TRUE) / sd(benchmark, na.rm = TRUE)  # Variability (Alpha)
  
  kge <- sqrt((r - 1)^2 + (Alpha - 1)^2 + (Beta)^2)  # Calculate KGE
  
  return(list(KGE = kge, r = r, Beta = Beta, Alpha = Alpha))  # Return all components
}
```

Then, we can calculate the metrics per model. To calculate the global performance, we average the metric values across all grid cells and years. We do this by taking a mean, weighted for the harvest area of the respective grid cell (HAWM). This means that grid cells where a lot of crops are produced, weight more in the calculation of the mean and thus have a higher influence than low-producing locations.

```{r, warning=FALSE, message = FALSE}
performance_metrics <- crop_data[["aggr"]] %>% 
  inner_join(general_sim_aggr %>% dplyr::select(lat, lon, year, difftrend_sim, model), by = c("lat", "lon", "year")) %>% 
   distinct() %>%
  drop_na(difftrend_obs, difftrend_sim) %>% # Remove rows with NAs in either column
  group_by(model, lat, lon) %>% 
  filter(sd(difftrend_obs) > 0, sd(difftrend_sim) > 0) %>%
  summarise(RMSE = rmse(difftrend_obs, difftrend_sim),
            Rsquared = cor(difftrend_obs, difftrend_sim)^ 2,
            KGE_components = calculate_kge(difftrend_obs, difftrend_sim),
            sd = sd(difftrend_sim),
           total_area = first(total_area)) %>% 
  mutate(
    KGE = KGE_components$KGE,
    r = KGE_components$r,
    Beta = KGE_components$Beta,
    Alpha = KGE_components$Alpha
  ) %>% dplyr::select(-KGE_components)  %>%  
  group_by(model) %>% 
  summarise(HAWM_RMSE = sum(RMSE*total_area)/sum(total_area),
            HAWM_Rsquared = sum(Rsquared*total_area)/sum(total_area),
            HAWM_KGE = sum(KGE*total_area)/sum(total_area),
            HAWM_Alpha = sum(Alpha*total_area)/sum(total_area),
            HAWM_beta = sum(Beta*total_area)/sum(total_area),
            HAWM_sd = sum(total_area * sd)/sum(total_area)) %>% 
  distinct()
```

Then we can plot these values in a heatmap table. Note that we have to normalize the values to create a common color scheme for lower to higher performance.

```{r, warning = FALSE, message = FALSE}
# Reshape and rename metric names for the table plot
performance_metrics_long <- performance_metrics %>%
  dplyr::select(model, HAWM_KGE, HAWM_sd, HAWM_RMSE, HAWM_Rsquared) %>%
  pivot_longer(cols = -model, names_to = "Metric", values_to = "Value") %>%
  mutate(Metric = recode(Metric,
                         "HAWM_KGE" = "KGE",
                         "HAWM_sd" = "Variability",
                         "HAWM_RMSE" = "RMSE",
                         "HAWM_Rsquared" = "Rsquared")) %>%
  group_by(Metric) %>%
  mutate(
    # Normalize value within each metric, reversing Rsquared for higher=better
    Normalized_Value = ifelse(
      Metric == "Rsquared",
      (Value - min(Value)) / (max(Value) - min(Value)),  # Rsquared higher is better
      (max(Value) - Value) / (max(Value) - min(Value))   # Others lower is better
    )
  ) %>%
  ungroup()

# Filter out 'Variability' for coloring purposes but keep for ordering
performance_metrics_long$fill_color <- ifelse(performance_metrics_long$Metric == "Variability", NA, performance_metrics_long$Normalized_Value)

# Create the table plot 
table_plot <- ggplot(performance_metrics_long, aes(x = Metric, y = reorder(model, Value), fill = fill_color)) +
  geom_tile(color = "black") +  # Outline tiles
  geom_text(aes(label = round(Value, 3)), color = "black", size = 3) +  # Add text labels
  geom_vline(xintercept = 3.5, color = "black", size = 1.5) +  # Thicker line at x=2.5 (between columns 2 and 3)
  scale_fill_viridis(
    discrete = FALSE,
    option = "viridis",
     na.value = "white" , # Color for 'Variability' set to white 
    name = "",  # Remove the legend title
    breaks = c(0, 1),  
    labels = c("Lower Performance", "Higher Performance")
  )  +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.text.y = element_text(size = 10),
    axis.title.x = element_blank(),  
  axis.title.y = element_blank(), 
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 7),
    legend.margin = margin(t = -10, unit = "pt"),  # Adjust top margin to move color bar closer
  legend.spacing.y = unit(0.1, "pt")            # Reduce spacing between legend and plot
  ) +
    labs(x = "", y = "", fill = "") +  # Add a label to the color bar
  guides(
    fill = guide_colorbar(
      title = NULL,
      barwidth = 16,
      barheight = 0.4,
    )
  ) 

table_plot

# Save the table plot
ggsave(file.path(base_path, "GGCMI-validation/results/main/fig1c_PerformanceTableHAWM.png"),
     table_plot, width = 6, height = 4, dpi = 300)
```

## Supplementary figure 1

In addition to the table heatmap plot above, we also create a version without weighting for the harvest area, treating each grid cell as equally important when taking the average.

```{r, warning = FALSE, message = FALSE}
performance_metrics <- crop_data[["aggr"]] %>% 
  inner_join(general_sim_aggr %>% dplyr::select(lat, lon, year, difftrend_sim, model), by = c("lat", "lon", "year")) %>% 
   distinct() %>%
  drop_na(difftrend_obs, difftrend_sim) %>% # Remove rows with NAs in either column
  group_by(model, lat, lon) %>% 
  filter(sd(difftrend_obs) > 0, sd(difftrend_sim) > 0) %>%
  summarise(RMSE = rmse(difftrend_obs, difftrend_sim),
            Rsquared = cor(difftrend_obs, difftrend_sim)^ 2,
            KGE_components = calculate_kge(difftrend_obs, difftrend_sim),
            sd = sd(difftrend_sim),
           total_area = first(total_area)) %>% 
  mutate(
    KGE = KGE_components$KGE,
    r = KGE_components$r,
    Beta = KGE_components$Beta,
    Alpha = KGE_components$Alpha
  ) %>% dplyr::select(-KGE_components)  %>%  
  group_by(model) %>% 
  summarise(mean_RMSE = median(RMSE),
            mean_Rsquared = median(Rsquared),
            mean_KGE = median(KGE),
            mean_Alpha = median(Alpha),
            mean_beta = median(Beta),
            mean_sd = median(sd)) %>% 
  distinct()

# Reshape and rename metric names for the table plot
performance_metrics_long <- performance_metrics %>%
  dplyr::select(model, mean_KGE, mean_sd, mean_RMSE, mean_Rsquared) %>%
  pivot_longer(cols = -model, names_to = "Metric", values_to = "Value") %>%
  mutate(Metric = recode(Metric,
                         "mean_KGE" = "KGE",
                         "mean_sd" = "Variability",
                         "mean_RMSE" = "RMSE",
                         "mean_Rsquared" = "Rsquared")) %>%
  group_by(Metric) %>%
  mutate(
    # Normalize value within each metric, reversing Rsquared for higher=better
    Normalized_Value = ifelse(
      Metric == "Rsquared",
      (Value - min(Value)) / (max(Value) - min(Value)),  # Rsquared higher is better
      (max(Value) - Value) / (max(Value) - min(Value))   # Others lower is better
    )
  ) %>%
  ungroup()

# Filter out 'Variability' for coloring purposes but keep for ordering
performance_metrics_long$fill_color <- ifelse(performance_metrics_long$Metric == "Variability", NA, performance_metrics_long$Normalized_Value)

# Create the table plot 
table_plot <- ggplot(performance_metrics_long, aes(x = Metric, y = reorder(model, Value), fill = fill_color)) +
  geom_tile(color = "black") +  # Outline tiles
  geom_text(aes(label = round(Value, 3)), color = "black", size = 3) +  # Add text labels
  geom_vline(xintercept = 3.5, color = "black", size = 1.5) +  # Thicker line at x=2.5 (between columns 2 and 3)
  scale_fill_viridis(
    discrete = FALSE,
    option = "viridis",
     na.value = "white" , # Color for 'Variability' set to white 
    name = "",  # Remove the legend title
    breaks = c(0, 1),  
    labels = c("Lower Performance", "Higher Performance")
  )  +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.text.y = element_text(size = 10),
    axis.title.x = element_blank(),  
  axis.title.y = element_blank(), 
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 7),
    legend.margin = margin(t = -10, unit = "pt"),  # Adjust top margin to move color bar closer
  legend.spacing.y = unit(0.1, "pt")            # Reduce spacing between legend and plot
  ) +
    labs(x = "", y = "", fill = "") +  # Add a label to the color bar
  guides(
    fill = guide_colorbar(
      title = NULL,
      barwidth = 16,
      barheight = 0.4,
    )
  ) 

table_plot

# Save the table plot
ggsave(file.path(base_path, "GGCMI-validation/results/appendix/suppfig1_TablePerformanceMean.png"),
     table_plot, width = 6, height = 4, dpi = 300)
```

## Supplementary figures 2-5

For the appendix we recreate the global maps for each crop as well. Note that for wheat we only plot winter wheat (wwh) and for rice we chose for the first growing season rice (ri1).

### Supplementary figure 2 – RICE (first growing season)

```{r, warning = FALSE, message = FALSE}
ens_map <- crop_data[["ri1"]] %>% 
  group_by(lat, lon) %>% 
 filter(sd(difftrend_obs) > 0, sd(difftrend_sim) > 0) %>% # only gridcells where the standard deviation is bigger than 0
    summarise(rsquared = cor(difftrend_obs, difftrend_sim) ^2,
             ctr = ctr) %>%
distinct()

map_rsquared <- base_map +
  geom_tile(data = ens_map, aes(x = lon, y = lat, color = rsquared), size = 0.2) +
  scale_color_viridis(name = expression(R^2), option = "viridis", limits = c(0, 0.75),
                      guide = guide_colorbar(barwidth = unit(18, "cm"),  # Increase colorbar width
                                            barheight = unit(1, "cm"),  # Set colorbar height
                                            title.position = "top",  # Position title above the colorbar
                                             title.vjust = 1,  # Add space between title and colorbar
                                              title.hjust = 0.5  # Center the title
                      ))+
  labs(x = "", y = "") +
  theme(legend.position = "bottom", legend.title = element_text(size = 17))

map_rsquared

# Save the figure
ggsave(file.path(base_path, "GGCMI-validation/results/main/suppfig2_GlobalPerformanceMap_ri1.png"), map_rsquared, width = 18, height = 8, dpi = 300)
```

### Supplementary figure 3 – (WINTER) WHEAT

```{r, warning = FALSE, message = FALSE}
ens_map <- crop_data[["wwh"]] %>% 
  group_by(lat, lon) %>% 
 filter(sd(difftrend_obs) > 0, sd(difftrend_sim) > 0) %>% # only gridcells where the standard deviation is bigger than 0
    summarise(rsquared = cor(difftrend_obs, difftrend_sim) ^2,
             ctr = ctr) %>%
distinct()

map_rsquared <- base_map +
  geom_tile(data = ens_map, aes(x = lon, y = lat, color = rsquared), size = 0.2) +
  scale_color_viridis(name = expression(R^2), option = "viridis", limits = c(0, 0.75),
                      guide = guide_colorbar(barwidth = unit(18, "cm"),  # Increase colorbar width
                                            barheight = unit(1, "cm"),  # Set colorbar height
                                            title.position = "top",  # Position title above the colorbar
                                             title.vjust = 1,  # Add space between title and colorbar
                                              title.hjust = 0.5  # Center the title
                      ))+
  labs(x = "", y = "") +
  theme(legend.position = "bottom", legend.title = element_text(size = 17))

map_rsquared

# Save the figure
ggsave(file.path(base_path, "GGCMI-validation/results/main/suppfig3_GlobalPerformanceMap_wwh.png"), map_rsquared, width = 18, height = 8, dpi = 300)
```

### Supplementary figure 4 – SOY

```{r, warning = FALSE, message = FALSE}
ens_map <- crop_data[["soy"]] %>% 
  group_by(lat, lon) %>% 
 filter(sd(difftrend_obs) > 0, sd(difftrend_sim) > 0) %>% # only gridcells where the standard deviation is bigger than 0
    summarise(rsquared = cor(difftrend_obs, difftrend_sim) ^2,
             ctr = ctr) %>%
distinct()

map <- base_map +
  geom_tile(data = ens_map, aes(x = lon, y = lat, color = rsquared), size = 0.2) +
  scale_color_viridis(name = expression(R^2), option = "viridis", limits = c(0, 0.75),
                      guide = guide_colorbar(barwidth = unit(18, "cm"),  # Increase colorbar width
                                            barheight = unit(1, "cm"),  # Set colorbar height
                                            title.position = "top",  # Position title above the colorbar
                                             title.vjust = 1,  # Add space between title and colorbar
                                              title.hjust = 0.5  # Center the title
                      ))+
  labs(x = "", y = "") +
  theme(legend.position = "bottom", legend.title = element_text(size = 17))

map

# Save the figure
ggsave(file.path(base_path, "GGCMI-validation/results/main/suppfig4_GlobalPerformanceMap_soy.png"), map, width = 18, height = 8, dpi = 300)
```

### Supplementary figure 5 – MAIZE

```{r, warning = FALSE, message = FALSE}
ens_map <- crop_data[["mai"]] %>% 
  group_by(lat, lon) %>% 
 filter(sd(difftrend_obs) > 0, sd(difftrend_sim) > 0) %>% # only gridcells where the standard deviation is bigger than 0
    summarise(rsquared = cor(difftrend_obs, difftrend_sim) ^2,
             ctr = ctr) %>%
distinct()

map <- base_map +
  geom_tile(data = ens_map, aes(x = lon, y = lat, color = rsquared), size = 0.2) +
  scale_color_viridis(name = expression(R^2), option = "viridis", limits = c(0, 0.75),
                      guide = guide_colorbar(barwidth = unit(18, "cm"),  # Increase colorbar width
                                            barheight = unit(1, "cm"),  # Set colorbar height
                                            title.position = "top",  # Position title above the colorbar
                                             title.vjust = 1,  # Add space between title and colorbar
                                              title.hjust = 0.5  # Center the title
                      ))+
  labs(x = "", y = "") +
  theme(legend.position = "bottom", legend.title = element_text(size = 17))

map

# Save the figure
ggsave(file.path(base_path, "GGCMI-validation/results/main/suppfig5_GlobalPerformanceMap_mai.png"), map, width = 18, height = 8, dpi = 300)
```

## Supplementary figures 6-10

In addition to the maps with general performance, we also plot the total production per grid cell observed in the benchmark data. We calculate the total production as the yield multiplied by the total area of the grid cell and then take the mean production value across the years.

### Supplementary figure 6 - aggregated crop production

```{r}
production <- crop_data[["aggr"]] %>% 
  mutate(production = yield * total_area * 1/10000 *1/1000) %>% 
  group_by(lat, lon) %>% 
  summarise(mean_production = mean(production, na.rm= TRUE))

map <- base_map +
  geom_tile(
    data = production,
    aes(x = lon, y = lat, color = mean_production),
    size = 0.2
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_color_viridis(
    name = "Aggregated crop production (kilotons per year)",
    option = "inferno",
    guide = guide_colorbar(
      barwidth = unit(18, "cm"),
      barheight = unit(1, "cm"),
      title.position = "top",
      title.vjust = 1,
      title.hjust = 0.5
    )
  )  +
  labs(x = "", y = "") +
  theme(legend.position = "bottom")

map

# Save the figure
ggsave(file.path(base_path, "/aggr_production.png"), map, width = 24, height = 12, dpi = 300)
```

### Supplementary figure 7 - RICE production

```{r}
production <- crop_data[["ri1"]] %>% 
  mutate(production = yield * total_area * 1/10000 *1/1000) %>% 
  group_by(lat, lon) %>% 
  summarise(mean_production = mean(production, na.rm= TRUE))

map <- base_map +
  geom_tile(
    data = production,
    aes(x = lon, y = lat, color = mean_production),
    size = 0.2
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_color_viridis(
    name = "Rice production (kilotons per year)",
    option = "inferno",
    guide = guide_colorbar(
      barwidth = unit(18, "cm"),
      barheight = unit(1, "cm"),
      title.position = "top",
      title.vjust = 1,
      title.hjust = 0.5
    )
  )  +
  labs(x = "", y = "") +
  theme(legend.position = "bottom")

map

# Save the figure
ggsave(file.path(base_path, "/ri1_production.png"), map, width = 24, height = 12, dpi = 300)
```

### Supplementary figure 8 - WHEAT production

```{r}
production <- crop_data[["wwh"]] %>% 
  mutate(production = yield * total_area  * 1/10000 *1/1000) %>% 
  group_by(lat, lon) %>% 
  summarise(mean_production = mean(production, na.rm= TRUE))

map <- base_map +
  geom_tile(
    data = production,
    aes(x = lon, y = lat, color = mean_production),
    size = 0.2
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_color_viridis(
    name = "Wheat production (kilotons per year)",
    option = "inferno",
    guide = guide_colorbar(
      barwidth = unit(18, "cm"),
      barheight = unit(1, "cm"),
      title.position = "top",
      title.vjust = 1,
      title.hjust = 0.5
    )
  )  +
  labs(x = "", y = "") +
  theme(legend.position = "bottom")

map

# Save the figure
ggsave(file.path(base_path, "/wwh_production.png"), map, width = 24, height = 12, dpi = 300)
```

### Supplementary figure 9 – SOY production

```{r}
production <- crop_data[["soy"]] %>% 
  mutate(production = yield * total_area  * 1/10000 *1/1000) %>% 
  group_by(lat, lon) %>% 
  summarise(mean_production = mean(production, na.rm= TRUE))

map <- base_map +
  geom_tile(
    data = production,
    aes(x = lon, y = lat, color = mean_production),
    size = 0.2
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_color_viridis(
    name = "Soy production (kilotons per year)",
    option = "inferno",
    guide = guide_colorbar(
      barwidth = unit(18, "cm"),
      barheight = unit(1, "cm"),
      title.position = "top",
      title.vjust = 1,
      title.hjust = 0.5
    )
  )  +
  labs(x = "", y = "") +
  theme(legend.position = "bottom")

map

# Save the figure
ggsave(file.path(base_path, "/soy_production.png"), map, width = 24, height = 12, dpi = 300)
```

### Supplementary figure 10 – MAIZE production

```{r}
production <- crop_data[["mai"]] %>% 
  mutate(production = yield * total_area  * 1/10000 *1/1000) %>% 
  group_by(lat, lon) %>% 
  summarise(mean_production = mean(production, na.rm= TRUE))

map <- base_map +
  geom_tile(
    data = production,
    aes(x = lon, y = lat, color = mean_production),
    size = 0.2
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_color_viridis(
    name = "Maize production (kilotons per year)",
    option = "inferno",
    guide = guide_colorbar(
      barwidth = unit(18, "cm"),
      barheight = unit(1, "cm"),
      title.position = "top",
      title.vjust = 1,
      title.hjust = 0.5
    )
  )  +
  labs(x = "", y = "") +
  theme(legend.position = "bottom")

map

# Save the figure
ggsave(file.path(base_path, "/mai_production.png"), map, width = 24, height = 12, dpi = 300)
```
